%----------------------------------
\chapter{Introduction}\label{sec:intro}
%----------------------------------

\input{TrickHLAIntro}

\section{Identification of Document}
This document describes the use of the
\TrickHLA\ developed for use in the Trick Simulation Environment.
This document adheres to the documentation standards defined in
NASA Software Engineering Requirements Standard \cite{NASA:SWE}.

\section{Scope of Document}
This document provides information on the use of the \TrickHLA.

\section{Purpose and Objectives of Document}
The purpose of this document is to describe how to incorporate the
\TrickHLA\ into a dynamic Trick simulation and used by other simulation models.

\section{Documentation Status and Schedule}
The information in this document is current with the \TrickHLAid\
implementation of the \TrickHLA. Updates will be kept current with
module changes.

\begin{tabular}{||l|l|l|} \hline
\documentHistory
\end{tabular}

\begin{tabular}{||l|l|l|} \hline
\DocumentChangeHistory
\end{tabular}

\section{Document Organization}
This document is organized into the following sections:

\begin{description}

\item[Chapter \ref{sec:intro}: Introduction] --
Identifies this document, defines the scope and purpose, present status,
and provides a description of each major section.

\item[Chapter \ref{sec:docs}: Related Documentation] --
Lists the related documentation that is applicable to this project.

\item[Chapter \ref{sec:preliminaries}: Preliminaries] --
Discusses some Trick, HLA and \TrickHLA\ concepts that are used
in the subsequent chapters.

\item[Chapter \ref{sec:simplesine-sim}: An Example \simplesine Simulation] --
Introduces the \simplesine model in the context of a non-HLA simulation.

\item[Chapter \ref{sec:hla-join}: Joining a Federation] --
Illustrates how to make a Trick simulation become an HLA federate
and join an HLA federation.

\item[Chapter \ref{sec:hla-pubsub}: Publishing and Subscribing] --
Shows how Trick simulations may publish and subscribe data.

\item[Chapter \ref{sec:hla-lag}: Lag Compensation] --
Demonstrates how HLA-induced lags due to time management may be removed
either by publihsers of the data or by subscribers.

\item[Chapter \ref{sec:hla-inter}: Sending and Receiving Interactions] --
Illustrates how to send and receive HLA interactions.

\item[Chapter \ref{sec:hla-own}: Ownership Transfer] --
Shows how to use the \TrickHLA\ mechanisms for {\em pushing} and {\em pulling}
ownership of HLA data.

\item[Chapter \ref{sec:hla-pack}: Data Encoding and Packing] --
Shows how simulation developers may pack and unpack HLA data.

\item[Chapter \ref{sec:hla-init}: Initialization] --
Explains how to use \TrickHLA\ to implement multiphase
federation initialization.

\item[Chapter \ref{sec:hla-objDel}: Object Deleted Notification] --
Explains how to use \TrickHLA\ to implement callbacks when an object was deleted
from the federation.

\item[Chapter \ref{sec:hla_fed_save_setup}: How to setup your trick federate to
initiate a federation save] --
Explains how to upgrade your trick model to utilize \TrickHLA\ routines to save
the federate.

\item[Chapter \ref{sec:hla_trick_fed_save}: Federate Save] --
Explains how to initiate a federation save from the trick model.

\item[Chapter \ref{sec:hla_trick_fed_restore}: Federate Restore] --
Explains how to initiate a federation restore from the trick model.

\item[Bibliography] --
Informational references associated with this document.

\item[Appendix \ref{sec:simplesine-files}: \simplesine Files] --
Provides listings of some of the \simplesine source files.

\item[Appendix \ref{sec:send-receive-inputs}: Interaction send/receive input files] --
Provides listings of the input files for the simulations that
send and receive HLA interactions.

\end{description}
