%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% File: TrickHLASpec-Intro.tex
%
% Purpose: TrickHLA Product Specification - Intro Chapter
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------
\chapter{Introduction}\label{sec:intro}
%----------------------------------

\input{TrickHLAIntro}

\section{Identification of Document}
This document describes the design of the \TrickHLA\ developed
for use in the Trick Simulation Environment.
This document adheres to the documentation standards
defined in NASA Software Engineering Requirements Standard \cite{NASA:SWE}.

\section{Scope of Document}
This document provides information on the algorithms used in and
the design of the source code associated with \TrickHLA.
This includes references to relevant Trick and HLA documents.

\section{Purpose and Objectives of Document}
The purpose of this document is to provide a thorough understanding of the
\TrickHLA\ product and its underlying architecture and design.

\section{Documentation Status and Schedule}
The information in this document is current with the \TrickHLAid\
implementation of \TrickHLA. 
Updates will be kept current with module changes.

\begin{tabular}{||l|l|l|} \hline
\documentHistory
\end{tabular}

\begin{tabular}{||l|l|l|} \hline
\DocumentChangeHistory
\end{tabular}

\section{Document Organization}
This document is organized into the following sections:

\begin{description}

\item[Chapter \ref{sec:intro}: Introduction] -
Identifies this document, defines the scope and purpose, present status,
and provides a description of each major section.

\item[Chapter \ref{sec:docs}: Related Documentation] -
Lists the related documentation that is applicable to this project.

\item[Chapter \ref{sec:architectural_design}: Architectural Design] -
Presents the top-level concepts behind \TrickHLA.

\item[Chapter \ref{sec:math_formulations}: Mathematical Formulations] -
Presents the mathematical formulations implemented by \TrickHLA.

\item[Chapter \ref{sec:interface_design}: Interface Design] -
Describes the main interfaces to the \TrickHLA\ model.

\item[Chapter \ref{sec:functional_design}: Functional Design] -
Presents the detailed design of the model.

\item[Chapter \ref{sec:versions}: Version Description] -
Identifies the configuration-managed items that comprise \TrickHLA.

\item[Bibliography] -
Informational references associated with this document.

\end{description}

